Predicting the Quality of CDR-H3 Antibody Loop Structural Models

### 2023-06-29 paper.tex
More rewriting and queries from ACRM

\usepackage{tabularx} 
\setlength\extrarowheight{2pt} % make the tables look less cramped
\usepackage{graphicx}
\usepackage{adjustbox}
\newcommand{\ca}{\mbox{C$\alpha$}}
\newcommand{\VH}{\mbox{V\kern-.1667em \lower.5ex\hbox{\scriptsize H}}}
\newcommand{\VL}{\mbox{V\kern-.1667em \lower.5ex\hbox{\scriptsize L}}}
\newcommand{\VHVL}{\mbox{\VH/\VL}}
\newcommand{\CH}[1]{\mbox{C\lower.5ex\hbox{\scriptsize H}#1}}
\newcommand{\CL}{\mbox{C\kern-.0833em \lower.5ex\hbox{\scriptsize L}}}
\newcommand{\FV}{\mbox{\it Fv}}
\newcommand{\Fab}{\mbox{\it Fab}}

@@

\begin{abstract}
  Therapeutic antibodies have shown an unprecedented pace of
  development and have brought new hope for the treatment of numerous
  diseases. Bioinformatics tools for modelling antibody structures
  have become invaluable for antibody engineering and the development
  of therapeutic antibodies. The antigen-binding site consists of six
  hypervariable loops, also known as the Complementary Determining
  Regions (CDRs), all of which can generally be modelled with high accuracy,
  except for CDR-H3, which has
  far greater length, sequence and structural variability, 
  making modelling it is considerably harder.

  Many approaches for antibody modelling, such as our
  abYmod software, have been developed. Although such efforts have
  improved prediction accuracy, the results for CDR-H3 are
  still inconsistent and require further improvement. Providing a
  confidence score for the structure predictions would aid in
  differentiating well-modelled structures from incorrectly modelled
  structures, giving the user a clearer understanding of the
  generated 3D-model reliability.

  We present a 3D-model quality predictor, combining domain knowledge
  with machine learning techniques to predict the accuracy of CDR-H3
  3D-models generated by antibody modelling software such as abYmod. The
  newly developed predictor scored a Matthews Correlation Coefficient
  of 0.99, and can thus be described as highly reliable. The predictor
  is made available at \url{http://www.bioinf.org.uk/abs/qualiloop/}
\end{abstract}

@@

Antibodies are highly specialized proteins of the immune system that
are produced in response to a foreign substance, called an antigen. A
mature antibody binds a specific antigen with high affinity and specificity. This sets them apart from other pharmaceuticals and makes them effective drugs with endless possibilities in application given their ability to target an immense variety of antigens.
In contrast with small drug molecules, antibodies can not
only bind pockets, but also flat, concave or convex
surfaces\cite{MacCallum1996}. Their unique characteristics have enabled 
researchers to develop efficient antibody drugs for treating cancers,
autoimmune disorders, infectious diseases and many more\cite{Lu2020}.
Four of the top 10 best-selling drugs in
2020 were monoclonal antibodies\cite{Urquhart2021}.

In order to design therapeutic antibodies rationally, knowledge of
their structure is essential. The acquired structural information can
be used to modify binding affinity to a target of interest,
predicting both the exact binding site and the antibody stability as
well as assessing immunogenicity\cite{Abhinandan2007}. As
experimental structure determination is costly and time
consuming, computational predictions of an antibody's structure are
used to streamline the process.

The variable
fragment (Fv) of an antibody contains the six complementarity determining regions
(CDRs, also known as hypervariable loops) which form the antigen binding site.
All except one of these loops can be clustered
into a limited number of `canonical structures'\cite{Al-Lazikani}. Therefore, modelling
these loops with adequate accuracy is commonly
achievable\cite{North2011}.  However, the CDR loop 3 of the 
heavy chain (CDR-H3) has a far greater sequence and length variability due to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained mostly unclassifiable\cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared with other protein loops\cite{Regep2017}. It was found
that over 75\% of CDR-H3 loops do not have a sub-{\AA}ngstr\"{o}m non-antibody
structural neighbour, while 30\% of CDR-H3 loops have a
completely unique structure, compared with under 3\% for all non-antibody
loops\cite{Regep2017}.

@@

According to the Kabat definition, the CDR-H3 loop is made up of the
residues H95--H102 (using the Kabat\cite{Kabat1992}, Chothia\cite{Al-Lazikani} or Martin\cite{Abhinandan2008} numbering schemes) in the heavy
chain, with a potential insertion site at position H100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long.

For shorter loops, a higher prediction accuracy can be achieved than
for longer CDR-H3 loops. This was also shown in the Antibody Modelling
Assessments (AMA), two blind contests that required researchers to
build three-dimensional structural models (`3D-models' --- used throughout this text to distinguid from machine-learning models `ML-models') from antibody sequences. The CDR-H3 loop
modelling quality achieved at the contests was, on average, much lower
for loops of longer lengths\cite{Almagro2011,Almagro2014}.

Several different approaches for generating 
3D-models from antibody sequences exist including
RosettaAntibody\cite{Sircar2009,Sivasubramanian2009}, 
ABodyBuilder\cite{Leem2016}, PIGSPro\cite{Lepore2017}, Lyra\cite{Klausen2015}, AbLooper\cite{Abanades2022} and our own abYmod (manuscript in preparation).
One of the most
used methods is RosettaAntibody, which implements template selection
and \emph{ab initio} CDR-H3 loop modelling using loop fragments and employing
specific angle restraints which bias the conformational space towards
so-called `kinked' loops\cite{Schoeder2021,Weitzner2017}. In
contrast, ABodyBuilder uses a database search algorithm
(FREAD\cite{Choi2010}) for CDR loop modelling.
abYmod \url{http://abymod.abysis.org/} utilizes extensive canonical
class definitions, \VHVL\ angle prediction and a large database of loop
structures (LoopDB) for CDR-H3 modelling.
Upon inputting an antibody sequence, abYmod assigns the
canonical class using a set of key residues\cite{Martin1996} and where
an exact match is not possible, a nearest class is identified.

abYmod selects light and heavy chain frameworks separately from PDB templates.
First these are selected on the basis of the number of matched
canonical classes and then on the basis of sequence identity.  The
\VHVL\ packing angle is currently selected from the parent that has
the best sequence identity over both chains, but an improved method is
currently in development. Any CDRs where there was no canonical match are then
grafted onto the framework. If there is no template of the correct
length for CDR-H3, the loop is built using LoopDB, a database of
CDR-H3-like loops from all proteins. Finally, Gromacs energy
minimization software is used to optimize the 3D-model. This method has
proven very effective and preliminary analysis suggests the method
achieves comparable results, or outperforms, other modelling software
(see Results).

Using these modelling methods, framework regions can
generally be predicted with great accuracy (better than 1\AA\
\ca~RMSD\cite{Almagro2014}), as one can often find a very similar
structure for the homology modelling process.  However, the CDR loops
are not as easily predicted due to their great diversity. If the
canonical conformation of CDR loops CDR-L1,L2,L3,H1,H2 can be identified, they too
can be modelled rather well, often within 1\AA\ \ca~RMSD, for CDR-H3 loops
the average is usually above 3\AA\cite{Almagro2011}. The average values are taken from the second antibody modelling assessment \cite{Almagro2014}.

ABodyBuilder is a modelling server that provides the user with a
confidence score for each region (e.g.\ CDR-H2) of the antibody
3D-model. The given score is the probability that a specific region
(e.g.\ CDR-H2) will be modelled within a specific RMSD
threshold\cite{Leem2016}. Thus, it can be used to obtain an expected
RMSD value for a given probability (default 75\%). For CDR-H3, this
score is calculated as a function of the loop length.  The confidence
scorer is described as robust, but less accurate in the case of CDR
loops owing to the lack of data\cite{Leem2016}. ABLooper also provides a
confidence metric for the CDR-H3 loop 3D-model, which is estimated by the
diversity of a set of predicted conformations for the same
loop\cite{Abanades2022}. However, it remains unclear whether a high
prediction diversity score points towards loops with multiple
conformations or a low quality 3D-model. Furthermore, it remains unclear
how well the generated diversity score reflects 3D-model
quality\cite{Abanades2022}.

Modelling the CDR-H3 loop is a hurdle for \emph{in silico} development of
therapeutic antibodies. Currently, there is no definite, reliable way
to determine how accurate a generated structural 3D-model is within the
CDR-H3 region. Therefore, we have produced a
user-friendly predictor of CDR-H3 3D-model quality. The predictor will give
the user an RMSD-range in {\AA}ngstr\"{o}ms, in which the generated 3D-model lies
with a high probability.
This information can guide the
user in the antibody engineering process. The user has the choice to
determine whether the 3D-model is to be used as is, or
whether the 3D-model should be re-worked.

@@

The predictive power of any machine learning model (`ML-model') is largely
dependent on the quality and size of the dataset on which it was trained. As
this is a non-linear, complex, multi-class classification problem, a
substantial amount of data was required. Thus, an extensive, verified
dataset of antibody structures called abYbank/AbDb\cite{Ferdous2018},
was utilised (1924 non-redundant
structures). The \ca\ root-mean-square deviation (RMSD) value
between the crystal structures
and modelled structures was calculated (see Methods) and
was used to classify 3D-models.

The classifier predicts whether a 3D-model has an RMSD of below 2\AA,
between 2--4\AA, or above 4\AA. These cutoff values were selected
based on the observation that abYmod generally produces a 3D-model with
RMSD below 4\AA. Incorrectly modelled structures (Figure~\ref{fig:AMA}) may
be identified by screening for structures estimated to have an RMSD
above 4\AA. If a very high-quality 3D-model is needed one should also
exclude 3D-models with RMSD above 2\AA.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{AMA.eps}
  \caption {RMSD values of the CDR-H3 loop for structures from the
    Antibody Modelling Assessment I (2011) and AMAII (2014). abYmod
    outperforms other modelling software in some instances, but also
    has much lower accuracy in few outlier cases. 
    Right: Ab01 is the rabbit antibody PDB:4MA3, which was
    excluded in the CDR-H3 modelling stage in AMAII due to
    difficulties modelling the overall structure previously. Ab01 is
    shown for the methods, where generated 3D-models were adequate for
    RMSD calculation.}
  \label{fig:AMA}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{logo.eps}
  \caption {Sequence Logo of the CDR-H3 loop sequence. Data on amino
    acid occurrence taken from \protect\url{http://abymod.abysis.org/} Visualized
    using Seq2Logo using Kabat Numbering}
  \label{fig:logo}
\end{figure}

After creating the feature dataset, it is pre-processed (cleaning,
scaling, encoding, see methods for details). Structures with a
resolution worse than 4\AA\ were removed.
Instances of antibody structures in our non-redundant dataset that matched in loop sequence (but the overall sequence was different) were
not removed as 3D-Models of some of these structures with the same loop
sequence differ significantly. 
The few large RMSD ranges may stem from
low resolution. For example, the loop sequence with the largest RMSD range 
has multiple structure files linked to it of varying quality, one of which has a resolution of only 3.00\AA. Residue differences near the loop may also
explain the conformational difference of the loop itself, 
even if the loop sequence does not differ. Some of these structures are
complexed while others are not, which may also affect the loop
structure (manuscript in preparation).

The target data (i.e.\ RMSD values) are transformed from numerical
values to nominal values so that they can be used for
classification. In order to define these nominal categories, the total
RMSD range must be divided into categories. This is done either by
creating uniform classes (e.g.\ 1--2\AA, 2--3\AA, etc.), the optimal size of which
must be determined, or by creating balanced classes. When
creating balanced classes, the upper and lower thresholds of a
category are chosen in such a way that each class contains an equal
number of instances. Initially, this approach was chosen to counteract the
skewness of the RMSD distribution. However, this was found to
have a negative effect on the final ML-model's predictive power, so
uniform classes were used.

The RMSD values are also transformed into a set of binary values according to a
list of RMSD thresholds (i.e. a 1 is assigned to above and 0 to below a given threshold). This is done so that binary ML-models can be
trained, which will predict the probability e.g.\ that the 3D-model's RMSD
is above 2\AA, 2.2\AA, 2.4\AA, and so on. Several of these binary classifiers form a first layer, each outputting a prediction on whether the model is above or below the respective threshold. The outcome values are then fed into a second-layer classifier (Figure~\ref{fig:flow}). The number of binary
classifiers incorporated into the first layer has a great effect on
the final ML-model, the general trend being that the more binary
classifiers are used, the better the nominal prediction.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{simple_flow.eps}
  \caption {Simplified pipeline for creating the final machine learning model that will predict 3D-model quality by giving its RMSD range.}
  \label{fig:flow}
\end{figure}

@@

As some features are in the form of amino acid names, these must be
encoded before they can be passed to a ML-model. The
encoding strategy often determines how efficiently the ML-model learns
and how much information can be extracted. Different strategies were
employed to represent protein sequences numerically, such as BLOSUM62\cite{Henikoff1992} and NLF\cite{Nanni2011} encoding (a non-linear Fisher transform of a large set of physicochemical properties). However, the simple four-feature physiochemical encoding strategy\cite{Abhinandan2010} was implemented for all ML-models, as it was found to be
the most effective, although PCA-3 BLOSUM62, a dimensionality-reduced
BLOSUM62 encoding method, achieved comparable results.
Feature selection was conducted to improve the ML-model's learning
capacity. A high-dimensional feature dataset bears the risk of
introducing excessive noise, facilitating ML-model over-fitting and can be
responsible for an overall decrease in ML-model performance and
stability. Each additional input feature forces the ML-model to
handle a more complex task, which consumes excess computational power
and time and provides more variables leading to over-fitting of the ML-model.

In order to
determine the most effective feature selection method, the
ML-model was trained on different feature sets selected using manual
selection as well as algorithmic selection strategies (see methods).
None of the
feature selection methods was a best fit for all ML-models. To create a ML-model implementing the encoding and feature selection strategies best suited for the specific types of ML-classifiers selected for the first-layer and top-layer model in combination, a number of different combinations were tested, summarized in Table~\ref{tab:results_table}. Additional ML-Models were discarded due to poor performance.

@@

The best ML-model, and its best hyperparameters, were then determined for
each binary RMSD target. The set of binary ML-models outputs a number of
predictions that give the likelihood of the 3D-model having an RMSD above
the threshold value of the respective ML-model. These predictions are
then added to the feature set, on which a top-layer classifier is then
trained (Figure~\ref{fig:method})\lilian{Incorrect reference}. Thus, a quasi-voting-system is incorporated into the final
classifier, in which a set of weaker classifiers vote on the ML-model
quality.

@@

\subsection{Machine Learning Model Performance}
The overall best final ML-model was composed of several different binary
classifiers, (Figure~\ref{fig:flow}) with an extreme gradient boosting (XGBoost) top-layer
nominal classifier. Features were selected using a recursive feature
elimination algorithm, through which the weakest feature is removed
recursively and the model performance is tested. In the final model, 9
features are included: tip\_pos, protrusion, length, total\_charge, nr\_charged,
identity, similarity, Hydropathy and Hydropathy\_diff
(See Table~\ref{tab:feature_table}).
\lilian{This appears to be the top level. What about a table showing the features used for each of the binary predictors and its MCC and anything else unique to that classifier (e.g. feature selection method)?}

A final MCC value of 0.99 could be achieved for an ML-model
using the abYmod log file as input as well as the loop 3D-model file
itself. This value slightly dropped to 0.92 if no such log file was
given. This is due to the fact that the template sequence
abYmod used to generate the 3D-model is unknown in the latter case
and therefore, sequence identity, similarity and hydropathy difference cannot be calculated. The energy is also not available.
\lilian{There is no option to upload the log file on the web site!}

The software
was tested on a test-set of antibody structures used in the 2014 and
2011 Antibody Modelling Assessments
\cite{Almagro2011,Almagro2014}. As the results depicted in
Figure~\ref{fig:AMA} show, abYmod generally achieves results similar to, or better than,
other modelling programs. However, some outliers with very high RMSD
values increase abYmod's RMSD average. The predictor in this work
would aim to identify such outlier 3D-models.

@@

\subsection{Computing}
All machine learning, feature selection and hyperparameter
optimization algorithms were implemented in Python. The Scikit-learn
library was used for training ML-models, the
Yellowbrick\cite{Bengfort2021} library was utilized for
visualization. All code is available at
\url{https://github.com/LilianDenzler/qualiloop}

@@

\subsection{Data Pre-Processing and Preparation}
Handling Null Values and Duplicates: The dataset containing target
RMSD values, and the calculated features was screened for null
values.
Rows that contained any null values were
removed from the dataset (11 rows in total).

@@

\subsubsection{Scaling}
Normalization and Standardization were tested as scaling methods. In normalization the range of the data is fixed between 0 and 1, while in Standardization the data is re-scaled to fit a Gaussian distribution. Both
approaches are greatly influenced by outliers and, ideally, such datapoints are
removed for optimal scaling. Here we define outliers as
datapoints that lie over 1.5 times the interquartile range (IQR) below
the first quartile or above the third quartile. The IQR is defined as
the range between quartile 1, i.e.\ the median of the lower half of the
data, and quartile 3, i.e.\ the median of the upper half of the
data. However, across all features there are a total of 632 outlier
values and removing such a large number of datapoints was not a viable
option. A robust scaler\cite{XXXX} \lilian{Citation} was also used, which uses statistics that are
robust to outliers. The median is set to zero and numerical features
are scaled to the interquartile range. \lilian{So which of these scaling methods did you actually use in the end?!}

>>>>

\usepackage{tabularx} 
\setlength\extrarowheight{2pt} % make the tables look less cramped
\usepackage{graphicx}
%\usepackage{adjustbox}
\newcommand{\ca}{\mbox{C$\alpha$}}
\newcommand{\VH}{\mbox{V\kern-.1667em \lower.5ex\hbox{\scriptsize H}}}
\newcommand{\VL}{\mbox{V\kern-.1667em \lower.5ex\hbox{\scriptsize L}}}
\newcommand{\VHVL}{\mbox{\VH/\VL}}
\newcommand{\CH}[1]{\mbox{C\lower.5ex\hbox{\scriptsize H}#1}}
\newcommand{\CL}{\mbox{C\kern-.0833em \lower.5ex\hbox{\scriptsize L}}}
\newcommand{\FV}{\mbox{\it Fv}}
\newcommand{\Fab}{\mbox{\it Fab}}
\newcommand{\carmsd}{\mbox{C$\alpha$-RMSD}}

@@

\begin{abstract}
  Therapeutic antibodies have shown an unprecedented pace of
  development and have brought new hope for the treatment of numerous
  diseases. Bioinformatics tools for modelling antibody structures
  have become invaluable for antibody engineering and the development
  of therapeutic antibodies. The antigen-binding site consists of six
  hypervariable loops, also known as the Complementary Determining
  Regions (CDRs), all of which can generally be modelled with high accuracy,
  except for CDR-H3, which has
  far greater length, sequence and structural variability, 
  making modelling it considerably harder.

  Many approaches for antibody modelling, such as our
  abYmod software, have been developed. Although such efforts have
  improved prediction accuracy, the results for CDR-H3 are
  still inconsistent and require further improvement. Providing a
  confidence score for the structure predictions would aid in
  differentiating well-modelled structures from incorrectly modelled
  structures, giving the user a clearer understanding of the
  reliability of the 3D-model.

  We present a 3D-model quality predictor, combining domain knowledge
  with machine learning techniques to predict the accuracy of CDR-H3
  3D-models generated by antibody modelling software such as abYmod. The
  newly developed predictor is highly reliabkle, with a Matthews Correlation Coefficient
  of 0.99. The predictor
  is made available at \url{http://www.bioinf.org.uk/abs/qualiloop/}
\end{abstract}

@@

Antibodies are highly specialized proteins of the immune system that
are produced in response to a foreign substance, known as an antigen. A
mature antibody binds a given antigen with high affinity and specificity. These characteristics allow them to be used as pharmaceuticals and make them effective drugs with endless possibilities in application given their ability to target an immense variety of antigens.
In contrast to small drug molecules, antibodies can not
only bind into pockets, but also flat, concave or convex
surfaces\cite{MacCallum1996}. Their unique characteristics have enabled 
researchers to develop efficient antibody drugs for treating cancers,
autoimmune disorders and infectious diseases amongst others\cite{Lu2020}.
Exclusing Covid, half of the top 10 best-selling drugs in
2022 were monoclonal antibodies\cite{Buntz:TopDrugs2022}.

In order to add a rational element to the design of therapeutic antibodies, knowledge of
their structure is essential. The acquired structural information can
be used to modify binding affinity to a target of interest,
predicting both the exact binding site and the antibody stability as
well as assessing immunogenicity\cite{Abhinandan2007}. As
experimental structure determination is costly and time
consuming, computational predictions of an antibody's structure are often
used to streamline the process.

The variable
fragment (Fv) of an antibody contains the six complementarity determining regions
(CDRs, also known as hypervariable loops) which form the antigen binding site.
All except one of these loops can be clustered
into a limited number of `canonical structures'\cite{Al-Lazikani}.
Since these have characteristic sequence motifs, modelling
these loops with good accuracy is commonly
achievable\cite{North2011}.  However, the third CDR of the 
heavy chain (CDR-H3) has a far greater sequence and length variability owing to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained mostly unclassifiable\cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared with other protein loops\cite{Regep2017}. It was found
that over 75\% of CDR-H3 loops do not have a sub-{\AA}ngstr\"{o}m non-antibody
structural neighbour, while 30\% of CDR-H3 loops have a
completely unique structure, compared with under 3\% for all non-antibody
loops\cite{Regep2017}.

@@

According to the Kabat definition, the CDR-H3 loop is made up of
residues H95--H102 (using the Kabat\cite{Kabat1992}, Chothia\cite{Al-Lazikani} or Martin\cite{Abhinandan2008} numbering schemes) in the heavy
chain, with a potential insertion site at position H100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long.

For shorter loops, a higher prediction accuracy can be achieved than
for longer CDR-H3 loops. This was also shown in the Antibody Modelling
Assessments (AMA), two blind contests that required researchers to
build three-dimensional structural models from antibody sequences.
Throughout the rest of this paper, we use the term `3D-models' to distinguish them from machine-learning models (`ML-models'). The CDR-H3 loop
model quality achieved at the contests was, on average, much lower
for loops of longer lengths\cite{Almagro2011,Almagro2014}.

Several different approaches for generating 
3D-models from antibody sequences exist including
RosettaAntibody\cite{Sircar2009,Sivasubramanian2009}, 
ABodyBuilder\cite{Leem2016}, PIGSPro\cite{Lepore2017},
Lyra\cite{Klausen2015}, AbLooper\cite{Abanades2022},
IgFold\cite{XXXX}, \lilian{Please add a reference}
and our own abYmod (Martin\etal, manuscript in preparation).
RosettaAntibody implements template selection
and \emph{ab initio} CDR-H3 loop modelling using loop fragments and employing
specific angle restraints which bias the conformational space towards
so-called `kinked' loops\cite{Schoeder2021,Weitzner2017}. In
contrast, ABodyBuilder uses a database search algorithm
(FREAD\cite{Choi2010}) for CDR loop modelling. Our own method,
abYmod, (Available at \url{http://abymod.abysis.org/}) utilizes extensive canonical
class definitions\cite{Martin1996}, \VHVL\ angle prediction and a large database of loop
structures from all PDB protein structures (LoopDB) for CDR-H3 modelling.

% abYmod selects light and heavy chain frameworks separately from PDB templates.
% First these are selected on the basis of the number of matched
% canonical classes and then on the basis of sequence identity.  The
% \VHVL\ packing angle is currently selected from the parent that has
% the best sequence identity over both chains, but an improved method is
% currently in development. Any CDRs where there was no canonical match are then
% grafted onto the framework. If there is no template of the correct
% length for CDR-H3, the loop is built using LoopDB, a database of
% CDR-H3-like loops from all proteins. Finally, Gromacs energy
% minimization software is used to optimize the 3D-model. This method has
% proven very effective and preliminary analysis suggests the method
% achieves comparable results, or outperforms, other modelling software
% (see Results).

Using these modelling methods, framework regions can
generally be predicted with high accuracy (better than 1\AA\
\carmsd\cite{Almagro2014}), as one can often find a very similar
structure for the homology modelling process.  However, the CDR loops
are not as easily predicted owing to their great diversity. If the
canonical conformation of CDR loops CDR-L1,L2,L3,H1,H2 can be identified, they too
can be modelled rather well, generally with better than 1\AA\ \carmsd, while for CDR-H3 loops,
the average is usually above 3\AA\cite{Almagro2011}. The average values are taken from the second antibody modelling assessment \cite{Almagro2014}.

ABodyBuilder is a modelling server that provides the user with a
confidence score for each region (e.g.\ CDR-H2) of the antibody
3D-model. The given score is the probability that a specific region
(e.g.\ CDR-H2) will be modelled within a specific \carmsd\
threshold\cite{Leem2016}. Thus, it can be used to obtain an expected
\carmsd\ value for a given probability (default 75\%). For CDR-H3, this
score is calculated as a function of the loop length.  The confidence
score is described as robust, but is less accurate in the case of CDR
loops owing to the lack of data\cite{Leem2016}. ABLooper also provides a
confidence metric for CDR-H3 3D-models, which is estimated by the
diversity of a set of predicted conformations for the same
loop\cite{Abanades2022}. While CDRs are generally not flexible,
16\% of CDR-H3s show a local conformational change of $>1.0\AA$
and 5\% $>2.0\AA$ on binding (Liu and Martin, manuscript in preparation).
It therefore remains unclear whether a high
prediction diversity score in CDR-H3 points towards loops with multiple
conformations or a low quality 3D-model. Furthermore, it remains unclear
how well the generated diversity score reflects 3D-model
quality\cite{Abanades2022}.

Modelling CDR-H3 is a hurdle for \emph{in silico} development of
therapeutic antibodies; currently, there is no definite, reliable way
to predict the accuracy of a 3D-model of 
CDR-H3. Therefore, we have produced a
user-friendly predictor of CDR-H3 3D-model quality. The predictor will give
the user an \carmsd-range in {\AA}ngstr\"{o}ms, in which the generated 3D-model lies
with a high probability.
The user has the choice of
determining whether the 3D-model should be used as is, or
whether the 3D-model should be re-worked.

@@

The predictive power of any machine learning model (`ML-model') is largely
dependent on the quality and size of the dataset on which it was trained. As
this is a non-linear, complex, multi-class classification problem, a
substantial amount of data was required. Thus, an extensive, verified
dataset of antibody structures called (AbDb\cite{Ferdous2018}),
was utilised containing 1924 non-redundant
structures. Models were built using abYmod and the \carmsd\ value
between the crystal structures
and modelled structures was calculated (see Methods) and
was used to classify the CDR-H3 3D-models.

The classifier predicts whether a 3D-model of CDR-H3 has an \carmsd\ of below 2\AA,
between 2--4\AA, or above 4\AA. These cutoff values were selected
based on the observation that abYmod generally produces a 3D-model with
\carmsd\ below 4\AA\  (Figure~\ref{fig:AMA}). 
If a high-quality 3D-model is needed, one should also
exclude 3D-models with \carmsd\ above 2\AA.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{AMA.eps}
  \caption {\carmsd\ values of the CDR-H3 loop for structures from the
    Antibody Modelling Assessment I (2011) and AMAII (2014). abYmod
    outperforms other modelling software in some instances, but also
    has much lower accuracy in few outlier cases. 
    Right: Ab01 is the rabbit antibody PDB:4MA3, which was
    excluded in the CDR-H3 modelling stage in AMAII owing to
    difficulties modelling the overall structure previously. Ab01 is
    shown for the methods, where generated 3D-models were adequate for
    RMSD calculation.}
  \label{fig:AMA}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{logo.eps}
  \caption {Sequence Logo of the CDR-H3 loop sequence. Data on amino
    acid occurrence taken from \protect\url{http://www.abysis.org/} and visualized
    using Seq2Logo using Kabat numbering.
    \lilian{Did you use all species for the frequencies or limit to, say, human?}}
  \label{fig:logo}
\end{figure}

After creating the feature dataset (see Methods, Table~\ref{tab:feature_table}), it is pre-processed (cleaning,
scaling, encoding. See Methods for details). Structures with a
resolution worse than 4\AA\ were removed.
Instances of antibody structures in our non-redundant dataset that matched in loop sequence (but where the overall sequence was different) were
not removed as 3D-Models of some of these structures with the same loop
sequence differ significantly. 
The few large \carmsd\ ranges may stem from
low resolution. For example, the loop sequence with the largest \carmsd\ range 
has multiple structure files linked to it of varying quality, one of which has a resolution of only 3.00\AA. Residue differences near the loop may also
explain the conformational difference of the loop itself, 
even if the sequence of the loop itself does not differ. Some of these structures are
complexed while others are not, which can also affect the loop
structure (Liu and Martin, manuscript in preparation).

The target data (i.e.\ \carmsd\ values) are transformed from numerical
values to nominal values so that they can be used for
classification. In order to define these nominal categories, the total
\carmsd\ range must be divided into categories. This is done either by
creating uniform classes (e.g.\ 1--2\AA, 2--3\AA, etc.), the optimal size of which
must be determined, or by creating balanced classes. When
creating balanced classes, the upper and lower thresholds of a
category are chosen in such a way that each class contains an equal
number of instances. Initially, this approach was chosen to counteract the
skewness of the \carmsd\ distribution. However, this was found to
have a negative effect on the final ML-model's predictive power, so
uniform classes were used.

The \carmsd\ values are also transformed into a set of binary values according to a
list of \carmsd\ thresholds (i.e. a 1 is assigned to above and 0 to below a given threshold). This is done so that binary ML-models can be
trained, which will predict the probability that the 3D-model's \carmsd\
is, for example, above or below 2\AA, 2.2\AA, 2.4\AA, and so on. Several of these binary classifiers form a first layer, each outputting a prediction on whether the model is above or below the respective threshold. The outcome values are then fed into a second-layer classifier (Figure~\ref{fig:flow}). The number of binary
classifiers incorporated into the first layer has a great effect on
the final ML-model, the general trend being that the more binary
classifiers are used, the better the nominal prediction.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{simple_flow.eps}
  \caption {Simplified pipeline for creating the final machine learning model that will predict 3D-model quality by giving its \carmsd\ range.}
  \label{fig:flow}
\end{figure}

@@

As some features are in the form of amino acid names, these must be
encoded before they can be passed to an ML-model. The
encoding strategy often determines how efficiently the ML-model learns
and how much information can be extracted. Different strategies were
employed to represent protein sequences numerically, such as
BLOSUM62\cite{Henikoff1992} and NLF\cite{Nanni2011} encoding (a
non-linear Fisher transform of a large set of physicochemical
properties). However, a simple four-feature physiochemical encoding
strategy\cite{Abhinandan2010} was found to be the most effective,
although PCA-3 BLOSUM62 (a dimensionality-reduced BLOSUM62 encoding
method) achieved comparable results. The simple physical encoding was
implemented for all ML-models.
Feature selection was conducted to improve the ML-model's learning
capacity. A high-dimensional feature dataset bears the risk of
introducing excessive noise, facilitating ML-model over-fitting and can be
responsible for an overall decrease in ML-model performance and
stability. Each additional input feature forces the ML-model to
handle a more complex task, which consumes excess computational power
and time and provides more variables leading to over-fitting of the ML-model.

In order to
determine the most effective feature selection method, the
ML-model was trained on different feature sets selected using manual
and algorithmic selection strategies (see Methods).
None of the
feature selection methods was a best fit for all ML-models. To create an ML-model implementing the encoding and feature selection strategies best suited for the specific types of ML-classifiers selected for the first-layer and top-layer models in combination, a number of different combinations were tested, summarized in Table~\ref{tab:results_table}. Additional ML-Models were discarded owing to poor performance.

@@

The best ML-model, together with its best hyperparameters, was then determined for
each binary \carmsd\ target. The set of binary ML-models outputs a number of
predictions that give the likelihood of the 3D-model having an \carmsd\ above
the threshold value of the respective ML-model. These predictions are
then added to the feature set, on which a top-layer classifier is then
trained (Figure~\ref{fig:method})\lilian{Incorrect reference}.
Thus, a quasi-voting-system is incorporated into the final
classifier, in which a set of weaker classifiers vote on the ML-model
quality.

@@

\subsection{Machine Learning Model Performance}
The overall best final ML-model was composed of several different binary
classifiers (Figure~\ref{fig:flow}), with an extreme gradient boosting (XGBoost) top-layer
nominal classifier. Features were selected using a recursive feature
elimination algorithm, through which the weakest feature is removed
recursively and the model performance is tested. In the final model, nine
features were included: tip\_pos, protrusion, length, total\_charge, nr\_charged,
identity, similarity, Hydropathy and Hydropathy\_diff
(See Methods, Table~\ref{tab:feature_table}).
\lilian{This appears to be the top level. What about a table showing the features used for each of the binary predictors and its MCC and anything else unique to that classifier (e.g. feature selection method)?}

A final MCC value of 0.99 could be achieved for an ML-model using the
3D-model together with the abYmod log file from which sequence
identity, similarity and hydropathy difference (all compared with the
parent structures) as well as the energy from energy minimization
could be included. For a predictor that doesn't incorporate these data
and required only the 3D-model (and could thus be used with different
modelling approaches), the MCC drops slightly to 0.92.
\lilian{What was the test set for these MCC values?}
\lilian{There is no option to upload the log file on the web site!}

The software
was tested on a test-set of antibody structures used in the 2014 and
2011 Antibody Modelling Assessments
\cite{Almagro2011,Almagro2014}. As the results depicted in
Figure~\ref{fig:AMA} show, abYmod generally achieves results similar to, or better than,
other modelling programs. However, some outliers with very high \carmsd\
values increase abYmod's \carmsd\ average. \highlight{The predictor in this work
  would aim to identify such outlier 3D-models.} \lilian{This needs rewording ---
  does the predictor actually succeed in doing so?!}

@@

\subsection{Computing}
All machine learning, feature selection and hyperparameter
optimization algorithms were implemented in Python. The Scikit-learn
library was used for training ML-models and the
Yellowbrick\cite{Bengfort2021} library was utilized for
visualization. All code is available at
\url{https://github.com/LilianDenzler/qualiloop}

@@

\subsection{Data Pre-Processing and Preparation}

\subsection{Handling Null Values}
The dataset containing target
\carmsd\ values and calculated features was screened for null
values which occur when a parameter cannot be calculated.
Rows that contained any null values were
removed from the dataset (11 rows in total). \lilian{From how many?}

@@

\subsubsection{Scaling}
Normalization and Standardization were tested as scaling methods. In normalization the range of the data is fixed between 0 and 1, while in Standardization the data are re-scaled to fit a Gaussian distribution. Both
approaches are greatly influenced by outliers and, ideally, such datapoints are
removed for optimal scaling. Here we define outliers as
datapoints that lie over 1.5 times the interquartile range (IQR) below
the first quartile or above the third quartile. The IQR is defined as
the range between quartile 1, i.e.\ the median of the lower half of the
data, and quartile 3, i.e.\ the median of the upper half of the
data. However, across all features there are a total of 632 outlier
values and removing such a large number of datapoints was not a viable
option. Consequently, a robust scaler\cite{XXXX} \lilian{Add Citation} was employed, which uses statistics that are
robust to outliers. The median is set to zero and numerical features
are scaled to the interquartile range. \lilian{I'm assimung this
  robust scaler was what was used for the final predictions?!}
